### 2017-02-24 IROS17/intro.tex

\section{Introduction}\label{sec:Intro}
Particle swarms propelled by a global field are common in applied mathematics, biology, and computer graphics. 
%If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
%\begin{align}
%[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
%\end{align}
%These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
% In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
%  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
% Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}-\citep{qiu2015magnetic}.
Someday large swarms of robots will be remotely guided
to assemble structures in parallel and 
 through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
 For each task, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
They require techniques to shape the swarm that can reliably exploit large populations despite severe underactuation ($2 \ll2n$).

Swarm shape control is necessary for navigation through narrow passages, for self-assembly, and for manipulating objects.
\begin{figure}
\centering
\vspace{1.5em}
\begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){n=2}\end{overpic}
\begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(10, 102){Optimal shortest path}\end{overpic}\\
\vspace{1em}
\begin{overpic}[width=0.43\columnwidth]{monalizaBegin.pdf}\put(30, 100){n=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
\begin{overpic}[width=0.38\columnwidth]{monalisa.pdf}\put(-3, 102){position control algorithm}\put(7,25){final configuration}\end{overpic}
\caption{\label{fig:IntroPic}
Dynamic algorithm to find the shortest path for two robots is designed. The two top pictures shows the initial position and goal position of the robots and the optimal shortest path by making use of wall friction when shared control input is applied. The two bottom pictures shows the n robot positioning using shared control input and boundary walls. 
} \vspace{-1em}
\end{figure}

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{subsec:WallFriction} introduces general models for boundary interaction.  We introduce algorithms with two orthogonal boundaries with high friction sufficient to arbitrarily position two robots in Sec.  \ref{sec:PostionControl2Robots}, and Sec.  \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
%Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.  \ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm correlation.
Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}.  We end with directions for future research in Sec.  \ref{sec:conclusion}.