### 2017-06-05 journalSwarmControl/coverletter/EditorNoteTROparticleSwarmManipulation.tex
final edits to paper (abstract too long)

Dear Editors,
 We are submitting our manuscript “Steering a Particle Swarm Using Global Inputs and Swarm Statistics”, to be considered for publishing in IEEE Transactions on Robotics (TRO). Authors are Shiva Shahrokhi, Lillian Lin, Chris Ertel, Mable Wan, and Aaron T. Becker.

This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a single journal paper. These papers are in IEEE International Conference on Robotics and Automation (ICRA) 2014 and IEEE International Conference on Intelligent Robots and Systems (IROS) 2015. The first conference paper covered the first three months of SwarmControl.net experiments [1], and the second one presented simulations of object manipulation [2]. We have incorporated numerous suggestions from external reviewers to increase the scope of this work. This paper presents three years of results from SwarmControl.net. For object manipulation, this paper presents robust new algorithms for manipulation, path planning, and obstacle avoidance, and a rich set of parameter sweeps over key variables. All hardware validation experiments are new, and feature a set of 100+ kilobots programmed to move toward a bright light source. Changing the location of the light source steers the kilobot swarm and uses them as a compliant manipulator.

We believe that the topic is a perfect fit for the scope of your journal and are hopeful that editors and reviewers share the opinion that this submission will be of interest to the readers of IEEE Transactions on Robotics.

With Best Regards,
Aaron T. Becker (on behalf of all authors)
713 743 6671, atbecker@uh.edu.
Assistant Professor, Electrical and Computer Engineering, University of Houston

[1] A. T. Becker, C. Ertel, and J. McLurkin, “Crowdsourcing swarm manipulation experiments: A massive online user study with large swarms of simple robots,” in IEEE International Conference on Robotics and Automation (ICRA), 2014, pp. 2825–2830.
 [2] S. Shahrokhi and A. T. Becker, “Stochastic swarm control with global inputs,” in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Sep. 2015, pp. 421–427.

Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion. 
Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
To identify the key parameters for particle manipulation, we used a collection of online games where players steer swarms of up to 500 particles to complete manipulation challenges. We recorded statistics from over ten thousand players. 
Inspired by techniques where human operators performed well, we investigate controllers that use only the mean and variance of the swarm. We prove the mean position is controllable and provide conditions under which variance is controllable.  We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the particle distribution. 
Finally, we employ these controllers as primitives for an object manipulation task and implement all controllers on 100 kilobots controlled by the direction of a global light source.