### 2017-07-27 ICRA-RA-L2018/experiment.tex
Final version of IROS is added. New paper started

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Demonstration}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\emph{Hardware Demonstration: Position Control of n Robots}\\
A hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used to implement  Alg. \ref{alg:PosControlNRobots}. 
 % Designs for each are available at \cite{Arun2016Thingiverse}. 
  The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}.
    Only discrete, 1 cm moves in the $x$ and $y$ directions are used. The goal configuration highlighted in the top right corner represents a `U' made of seven particles. The dark red configuration is the current position of the particles. 
Due to the discretized movements allowed by the boundary, drift moves follow a 1 cm square.  Free particles return to their start positions but particles on the boundary to move laterally, generating a net sliding motion.

Fig. \ref{fig:construction2d} follows the motion of the particles through iterations  $k$=1, 2, and 7. All particles receive the same control inputs, but boundary interactions break the control symmetry.  Particles reach their goal positions in a first-in, first-out arrangement beginning with the bottom-left particle from the staging zone occupying the top-right position of the build zone.