### 2018-05-31 journalWallFriction/SwarmShapeControl.tex
updated pdf info, moved submitted version to folder.

%  compress using: gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=foo-ShapingSwarmCompressed.pdf ShapingSwarmFrictionSharedInput.pdf

%\documentclass[conference]{IEEEtran}
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts% This command is only needed if 
                                                          % you want to use the \thanks command
%\overrideIEEEmargins                                      % Needed to meet printer requirements.
\usepackage{times}

\makeatletter 
\let\NAT@parse\undefined
\makeatother

% numbers option provides compact numerical references in the text. 
%\usepackage[numbers]{natbib}
\usepackage{multicol}
%\usepackage[bookmarks=true]{hyperref}

\usepackage{bbm}
\usepackage{calc}
\usepackage{url}
\usepackage{transparent}
\usepackage[bookmarks=true,hidelinks]{hyperref}

\usepackage{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{bm}
\usepackage{amssymb}
\usepackage{rotating}

\usepackage{chngcntr}
\counterwithin{paragraph}{subsection} % makes paragraph depend on subsection

%\usepackage{xfrac}
\usepackage{nicefrac}
\usepackage{cite}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage[usenames, dvipsnames]{color}
\usepackage{colortbl}
%\usepackage{caption}

%\usepackage{wrapfig}
\usepackage{overpic}
%\usepackage{subfigure}
%\usepackage{textcomp}
\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
\usepackage{breqn} %for breaking equations automatically
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
%\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{todonotes}
\usepackage{authblk}
\usepackage[per-mode=symbol, detect-weight=true, binary-units=true]{siunitx}
\usepackage{rotating}
\usepackage{tikz}
\usepgflibrary{arrows}% for more options on arrows
%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
% uncomment this to hide all red todos
%\renewcommand{\todo}{}

%% ABBREVIATIONS
\newcommand{\qstart}{q_{\text{start}}}

%% MACROS

\providecommand{\abs}[1]{\left\lvert#1\right\rvert}
\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
\providecommand{\normn}[2]{\left\lVert#1\right\rVert_#2}
\providecommand{\dualnorm}[1]{\norm{#1}_\ast}
\providecommand{\dualnormn}[2]{\norm{#1}_{#2\ast}}
\providecommand{\set}[1]{\lbrace\,#1\,\rbrace}
\providecommand{\cset}[2]{\lbrace\,{#1}\nobreak\mid\nobreak{#2}\,\rbrace}
\providecommand{\lscal}{<}
\providecommand{\gscal}{>}
\providecommand{\lvect}{\prec}
\providecommand{\gvect}{\succ}
\providecommand{\leqscal}{\leq}
\providecommand{\geqscal}{\geq}
\providecommand{\leqvect}{\preceq}
\providecommand{\geqvect}{\succeq}
\providecommand{\onevect}{\mathbf{1}}
\providecommand{\zerovect}{\mathbf{0}}
\providecommand{\field}[1]{\mathbb{#1}}
\providecommand{\C}{\field{C}}
\providecommand{\R}{\field{R}}
\newcommand{\Cspace}{\mathcal{Q}}
\newcommand{\Uspace}{\mathcal{U}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\Hcal}{$\mathcal{H}$}
\providecommand{\Vcal}{$\mathcal{V}$}
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\homog}{homog}
\DeclareMathOperator{\domain}{dom}
\DeclareMathOperator{\range}{range}
\DeclareMathOperator{\sign}{sgn}
\providecommand{\polar}{\triangle}
\providecommand{\ainner}{\underline{a}}
\providecommand{\aouter}{\overline{a}}
\providecommand{\binner}{\underline{b}}
\providecommand{\bouter}{\overline{b}}
\newcommand{\D}{\nobreakdash-\textsc{d}}
%\newcommand{\Fspace}{\mathcal{F}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\free}{\text{\{}\mathsf{free}\text{\}}}
\providecommand{\iff}{\Leftrightarrow}
\providecommand{\subinner}[1]{#1_{\text{inner}}}
\providecommand{\subouter}[1]{#1_{\text{outer}}}
\providecommand{\Ppoly}{\mathcal{X}}
\providecommand{\Pproj}{\mathcal{Y}}
\providecommand{\Pinner}{\subinner{\Pproj}}
\providecommand{\Pouter}{\subouter{\Pproj}}
\DeclareMathOperator{\argmax}{arg\,max}
\providecommand{\Aineq}{B}
\providecommand{\Aeq}{A}
\providecommand{\bineq}{u}
\providecommand{\beq}{t}
\DeclareMathOperator{\area}{area}
\newcommand{\contact}[1]{\Cspace_{#1}}
\newcommand{\feasible}[1]{\Fspace_{#1}}
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\figwid}{0.22\columnwidth}
\newcommand{\TRUE}{\textbf{true}}
\newcommand{\FALSE}{\textbf{false}}
\newcommand{\xupdownarrow}[1]{%
  {\left\updownarrow\vbox to #1{}\right.\kern-\nulldelimiterspace}
}
%\newcommand{\xleftrightarrow}[1]{%
%  {\left\leftrightarrow\vbox to #1{}\right.\kern-\nulldelimiterspace}
%}
\DeclareMathOperator{\atan2}{atan2}
\allowdisplaybreaks

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}

\pdfinfo{
   /Author (Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker)
   /Title  (Exploiting Non-Slip Wall Contacts  to Position Two Particles Using The Same Control Input)
   /CreationDate (D:20160129120000)
   /Subject (Simple Robots)
   /Keywords (Robots;Uniform Control Inputs, Underactuated Robots,Motion Control,Path Planning for Multiple Mobile Robot Systems)
}

% paper title
\title{\LARGE \bf Exploiting Non-Slip Wall Contacts \\  to Position Two Particles Using The Same Control Input}

% You will get a Paper-ID when submitting a pdf file to the conference system
\author{Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker% <-this % stops a space
\thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]} and \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}.}% <-this % stops a space
\thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2,atbecker\}@uh.edu}}%
}
%\affil{Department of Electrical and Computer Engineering, \\
% University of Houston, Houston, TX 77204-4005 USA\\
% {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}}
%\thanks{S. Shahrokhi, A. Mahadev and  A. Becker are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204-4005 USA {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}
%}
%} %\end thanks

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\begin{abstract}

Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
 This paper investigates particle control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
In previous work, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
Because in vivo environments are usually not square, this paper extends the previous work to all convex workspaces and to 3D positioning. 
This paper also implements the algorithms using a hardware setup inspired by the gastrointestinal tract.

% KEYWORDS:   
%Underactuated Robots
%Motion Control
%Path Planning for Multiple Mobile Robot Systems
%	
%This is an extension of work presented in a preliminary conference paper:
%S. Shahrokhi, A. Mahadev, and A. T. Becker, ÒAlgorithms for shaping a particle swarm with a shared input by exploiting non-slip wall contacts,Ó in Intelligent Robots and Systems (IROS), 2017 IEEE/RSJ International Conference on, 2017.
\end{abstract}

\IEEEpeerreviewmaketitle

%%%%%%%%%%%%%%%
\input{intro}
%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
\input{relatedWork}
%%%%%%%%%%%%%%%
\input{theory}
\input{theory2robAlgs}
%\input{theorynrobAlgs}
%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
\input{simulation}
%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
\input{experiment}
%%%%%%%%%%%%%%%
\input{conclusion}
%%%%%%%%%%%%%%%
%\section*{Acknowledgments}
%%%%%%%%%%%%%%%
%% Use plainnat to work nicely with natbib. 
{%\footnotesize
%\bibliographystyle{plainnat}
%\bibliographystyle{SageH}
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,SwarmShapeControl}
}

% Uncomment to add appendix:
%\input{appendix}

\end{document}