### 2018-07-12 Qualifying Exam/chap/abs.tex

% abstract for thesis

%This is the abstract.  It has a limit of 150 words for a masters thesis and 350 words for a PhD dissertation.

Micro-robots are small enough to move through the passageways of the body, therefore they are suited for targeted drug delivery and micro-scale manufacturing. Due to their small size, a single robot does not have enough force to deliver payloads, and it is prohibitively difficult to have onboard computation. Therefore, these robots are usually controlled by global inputs such as a uniform external magnetic field. This thesis presents controllers and algorithms for steering such an under-actuated swarm. This work first proves that the mean position of the swarm is controllable, and shows how an obstacle can make the variance controllable. Then it derives automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the swarm distribution. Finally, this work uses friction with boundary walls to break the symmetry caused by the global input and uses it to steer two particles to arbitrary positions.