### 2018-09-06 ICRA19/intro.tex
Some more todos for Shiva & Haoran

\section{Introduction}\label{sec:Intro}

   \textcolor{red}{TODO: lets do some experiments with salt in 3D printed shapes (or laser cut). Salt is micro scale, and roughly uniform (I think) we can get different magnitudes of global force by varying the tilt angle.}

   \textcolor{red}{TODO: this text is from a previous paper}

Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
%If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
%\begin{align}
%[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
%\end{align}
%These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
% In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
%  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
% Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
%As a current example, micro- and nano-robots can be manufactured in large numbers, see~\cite{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
%Someday large swarms of robots will be remotely guided
%to assemble structures in parallel and 
% through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
% For each task, large numbers of micro robots are required to deliver sufficient payloads, but

The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
 The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
 Techniques are needed that can handle this underactuation.

 Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
 In previous work, we provided position control algorithms that only require non-slip wall contacts to position two robots \cite{shahrokhi2017}. 
We assumed that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall. This work analyses interactions between the boundary and the particle swarm to control shape of the swarm. In \cite{shahrokhi2018TRO} we showed how to control swarm statistics including mean position and variance. This work extends that work to covariance control.

 \begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =\figwid]{leadfigure.pdf}%\put(1,55){a)}
\end{overpic}
\caption{\label{fig:leadfigure} 
      \textcolor{blue}{TODO: we need a lead figure in column 1:   Combine kilobots from Fig 3 and 4 with photos of angle of repose by Haoran of iron with 4 rod angles. }
}
\end{figure}

%    \begin{figure}
%    \centering
%    \vspace{1.5em}
%    %\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
%    \begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
%    \begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
%    \vspace{1em}
%    \caption{\label{fig:IntroPic}
%    Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
%    } \vspace{-1em}
%    \end{figure}
%\todo{add the picture of magnetic setup}

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{sec:angle} introduces angle of repose, a parameter of the particle swarm that can be used for shape control. 
Section  \ref{sec:covControl} describes implementations of the algorithms of covariance control in simulation and hardware experiments. 
 % Sec.  \ref{sec:expResults} 
 We end with directions for future research in Sec.  \ref{sec:conclusion}.