### 2019-01-03 journalWallFriction/SwarmShapeControl.tex
added stubs for the final journal version.  Still need pictures

%  compress using: gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=foo-ShapingSwarmCompressed.pdf ShapingSwarmFrictionSharedInput.pdf

% And here is the link to the videos and summary and read me:
%  https://www.dropbox.com/home/BoundaryInteractionVideos
% difference between two pdf files: https://draftable.com/compare
%\documentclass[conference]{IEEEtran}
\documentclass[journal]{IEEEtran}
\IEEEoverridecommandlockouts% This command is only needed if 
                                                          % you want to use the \thanks command
%\overrideIEEEmargins                                      % Needed to meet printer requirements.
\usepackage{times}

\makeatletter 
\let\NAT@parse\undefined
\makeatother

% numbers option provides compact numerical references in the text. 
%\usepackage[numbers]{natbib}
\usepackage{multicol}
%\usepackage[bookmarks=true]{hyperref}

\usepackage{bbm}
\usepackage{calc}
\usepackage{url}
\usepackage{transparent}
\usepackage[bookmarks=true,hidelinks]{hyperref}

\usepackage{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{bm}
\usepackage{amssymb}
\usepackage{rotating}

\usepackage{chngcntr}
\counterwithin{paragraph}{subsection} % makes paragraph depend on subsection

%\usepackage{xfrac}
\usepackage{nicefrac}
\usepackage{cite}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage[usenames, dvipsnames]{color}
\usepackage{colortbl}
%\usepackage{caption}

%\usepackage{wrapfig}
\usepackage{overpic}
%\usepackage{subfigure}
%\usepackage{textcomp}
\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
\usepackage{breqn} %for breaking equations automatically
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
%\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{todonotes}
\usepackage{authblk}
\usepackage[per-mode=symbol, detect-weight=true, binary-units=true]{siunitx}
\usepackage{rotating}
\usepackage{tikz}
\usepgflibrary{arrows}% for more options on arrows
%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
% uncomment this to hide all red todos
%\renewcommand{\todo}{}

%% ABBREVIATIONS
\newcommand{\qstart}{q_{\text{start}}}

%% MACROS

\providecommand{\abs}[1]{\left\lvert#1\right\rvert}
\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
\providecommand{\normn}[2]{\left\lVert#1\right\rVert_#2}
\providecommand{\dualnorm}[1]{\norm{#1}_\ast}
\providecommand{\dualnormn}[2]{\norm{#1}_{#2\ast}}
\providecommand{\set}[1]{\lbrace\,#1\,\rbrace}
\providecommand{\cset}[2]{\lbrace\,{#1}\nobreak\mid\nobreak{#2}\,\rbrace}
\providecommand{\lscal}{<}
\providecommand{\gscal}{>}
\providecommand{\lvect}{\prec}
\providecommand{\gvect}{\succ}
\providecommand{\leqscal}{\leq}
\providecommand{\geqscal}{\geq}
\providecommand{\leqvect}{\preceq}
\providecommand{\geqvect}{\succeq}
\providecommand{\onevect}{\mathbf{1}}
\providecommand{\zerovect}{\mathbf{0}}
\providecommand{\field}[1]{\mathbb{#1}}
\providecommand{\C}{\field{C}}
\providecommand{\R}{\field{R}}
\newcommand{\Cspace}{\mathcal{Q}}
\newcommand{\Uspace}{\mathcal{U}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\Hcal}{$\mathcal{H}$}
\providecommand{\Vcal}{$\mathcal{V}$}
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\homog}{homog}
\DeclareMathOperator{\domain}{dom}
\DeclareMathOperator{\range}{range}
\DeclareMathOperator{\sign}{sgn}
\providecommand{\polar}{\triangle}
\providecommand{\ainner}{\underline{a}}
\providecommand{\aouter}{\overline{a}}
\providecommand{\binner}{\underline{b}}
\providecommand{\bouter}{\overline{b}}
\newcommand{\D}{\nobreakdash-\textsc{d}}
%\newcommand{\Fspace}{\mathcal{F}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\free}{\text{\{}\mathsf{free}\text{\}}}
\providecommand{\iff}{\Leftrightarrow}
\providecommand{\subinner}[1]{#1_{\text{inner}}}
\providecommand{\subouter}[1]{#1_{\text{outer}}}
\providecommand{\Ppoly}{\mathcal{X}}
\providecommand{\Pproj}{\mathcal{Y}}
\providecommand{\Pinner}{\subinner{\Pproj}}
\providecommand{\Pouter}{\subouter{\Pproj}}
\DeclareMathOperator{\argmax}{arg\,max}
\providecommand{\Aineq}{B}
\providecommand{\Aeq}{A}
\providecommand{\bineq}{u}
\providecommand{\beq}{t}
\DeclareMathOperator{\area}{area}
\newcommand{\contact}[1]{\Cspace_{#1}}
\newcommand{\feasible}[1]{\Fspace_{#1}}
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\figwid}{0.22\columnwidth}
\newcommand{\TRUE}{\textbf{true}}
\newcommand{\FALSE}{\textbf{false}}
\newcommand{\xupdownarrow}[1]{%
  {\left\updownarrow\vbox to #1{}\right.\kern-\nulldelimiterspace}
}
%\newcommand{\xleftrightarrow}[1]{%
%  {\left\leftrightarrow\vbox to #1{}\right.\kern-\nulldelimiterspace}
%}
\DeclareMathOperator{\atan2}{atan2}
\allowdisplaybreaks

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\definecolor{myMagenta}{RGB}{233,63,246}
\definecolor{myDarkGreen}{RGB}{97,173,89}

\newcommand{\revision}[1]{\textcolor{black}{#1}}  %changed to black for final

\pdfinfo{
   /Author (Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker)
   /Title  (Exploiting Non-Slip Wall Contacts  to Position Two Particles Using The Same Control Input)
   /CreationDate (D:20160129120000)
   /Subject (Simple Robots)
   /Keywords (Robots;Uniform Control Inputs, Underactuated Robots,Motion Control,Path Planning for Multiple Mobile Robot Systems)
}

% paper title
\title{Exploiting Non-Slip Wall Contacts  to Position\\ Two Particles Using The Same Control Input}

% You will get a Paper-ID when submitting a pdf file to the conference system
\author{Shiva Shahrokhi,~\IEEEmembership{Member,~IEEE,} Jingang Shi,  Benedict Isichei, and Aaron T. Becker,~\IEEEmembership{Senior Member,~IEEE,}% <-this % stops a space
\thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}, \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}, and  \href{https://nsf.gov/awardsearch/showAward?AWD_ID=1646566}{[CNS-1646566]}.}% <-this % stops a space
\thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2,atbecker\}@uh.edu}}%
\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
%\affil{Department of Electrical and Computer Engineering, \\
% University of Houston, Houston, TX 77204-4005 USA\\
% {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}}
%\thanks{S. Shahrokhi, A. Mahadev and  A. Becker are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204-4005 USA {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}
%}
%} %\end thanks
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shahrokhi \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\begin{abstract}

Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
For example, magnetic actuation of particles has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
 This paper investigates position control of particles using uniform forces (the same force is applied everywhere in the workspace).
Given a controllable field that can generate bidirectional forces in three orthogonal directions, steering one particle in 3D is trivial. 
Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
However, the walls of \emph{in vivo} and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
In previous work, we showed that the individual 2D position of two particles is controllable using global inputs in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
Because \emph{in vivo} environments are usually not square, this paper extends the previous work to all convex workspaces, \revision{ and shows how this could be extended} to 3D positioning \revision{of neutrally buoyant particles.} 
We investigate analytically an idealized variant of this problem with non-slip boundaries and control inputs that are applied uniformly to all particles in the workspace.
This paper also implements the algorithms \revision{in 2D} using a hardware setup inspired by the gastrointestinal tract.

% SHORTER ABSTRACT:
%Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots. This paper investigates position control of particles using uniform forces (the same force is applied everywhere in the workspace). Given a controllable field that can generate bidirectional forces in three orthogonal directions, steering one particle in 3D is trivial. Adding additional particles to steer makes the system underactuated because there are more states than control inputs. However, the walls of in vivo and artificial environments often have surface roughness such that particles do not move unless actuation pulls them away from the wall. This paper shows that the individual 2D position of two particles is controllable using global inputs in all convex workspaces with non-slip wall contact, and shows how this could be extended to 3D positioning of neutrally buoyant particles. We investigate analytically an idealized variant of this problem with non-slip boundaries and control inputs that apply uniformly to all particles in the workspace. This paper implements the algorithms in 2D using a hardware setup inspired by the gastrointestinal tract.

% KEYWORDS:   
% Underactuated Robots
% Motion Control
% Path Planning for Multiple Mobile Robot Systems
% Configuration Space
%	
%This is an extension of work presented in a preliminary conference paper:
%S. Shahrokhi, A. Mahadev, and A. T. Becker, ÒAlgorithms for shaping a particle swarm with a shared input by exploiting non-slip wall contacts,Ó in Intelligent Robots and Systems (IROS), 2017 IEEE/RSJ International Conference on, 2017.
\end{abstract}
\begin{IEEEkeywords}
Underactuated Robots, Motion Control, Path Planning for Multiple Robot Systems, Configuration Space.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

%%%%%%%%%%%%%%%
\input{intro}
%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
\input{relatedWork}
%%%%%%%%%%%%%%%
\input{theory}
\input{theory2robAlgs}
%\input{theorynrobAlgs}
%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
\input{simulation}
%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
\input{experiment}
%%%%%%%%%%%%%%%
\input{conclusion}
%%%%%%%%%%%%%%%
%\section*{Acknowledgments}
%%%%%%%%%%%%%%%
%% Use plainnat to work nicely with natbib. 
{%\footnotesize
%\bibliographystyle{plainnat}
%\bibliographystyle{SageH}
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,SwarmShapeControl}
}
%Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker
%
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{shiva.jpg}}]{Shiva Shahrokhi}(S`14) received M.S. and PhD degrees in the Electrical and Computer Engineering Department at the University of Houston in 2018.  
She received the B.S. degree in Computer Engineering from Iran University of Science and Technology,
Tehran, Iran. She is involved in studies of swarm manipulation and control for robotics applications.
These include steering massively under-actuated mobile robotic ensembles, swarm shape control, and machine learning.
\end{IEEEbiography}
%
\begin{IEEEbiography}{Jingang Shi}
Biography text here.
\end{IEEEbiography}
%
\begin{IEEEbiography}{Benedict Isichei}
received an M.S. degree in the Electrical and Computer Engineering Department at the University of Houston in 2018.
\end{IEEEbiography}
%
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{DrB.jpg}}]{Aaron T. Becker}(S`06-M`12-SM`17) received M.S. and PhD degrees in electrical and computer engineering from the University of Illinois at Urbana-Champaign, in 2008 and 2012. He was a post doctoral research scholar at the Multi-Robot Systems Lab at Rice University and at Boston Children's Hospital \& Harvard Medical School, before joining the Electrical and Computer Engineering Department at the University of Houston as an Assistant Professor in 2014. He won the best paper at IROS 2013 and received the NSF CAREER award in 2016. %He is active in developing control laws and algorithms for robot swarms.
\end{IEEEbiography}

\end{document}